---
title: "Introductory analysis"
author: "Edoardo Filippi"
execute:
  echo: true
  warning: false
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    number_sections: true
    theme: cosmo

---
How this works: 

With these Parameters you can control the analysis

The workflow is supposed to be: run it once setting the filtering that you want and the contrasts that you need, the objects will be saved in the "analysis_name" sub-folder in the analysis results folder. When you want to run it again with the same parameters and, for example, create a report, you can set the load_dds_objects to true, they will be loaded and unnecessary operations skipped.

```{r params, eval = TRUE}
params <- list(
  "workflow" = "dds_gse",
  "subset_gse" = TRUE,
  "load_gse" = TRUE,
  "load_dds_objects" = FALSE,
  "save_dds_objects" = TRUE,
  "analysis_name" = "",
  "sample_metadata_file" = "",
  "create_annotation_df" = TRUE,
  "species" = "Hs" # or "Mm"
)
```

# setup

This is the function to modify when you want to filter the object in a specific way before beginning. 
All the subsequent analyses will be carried out starting from the samples present in this object.

```{r subsetting-object, eval = TRUE, include = TRUE}
subset_gse <- function(gse_to_subset) {

  # gse_to_subset <- gse_to_subset[, colData(gse_to_subset)$group != "WT"]
  # gse_to_subset <- gse_to_subset[, colData(gse_to_subset)$group %in% c("WT", "KO")]

  return(gse_to_subset)
}
```


```{r setup, include=FALSE, cache=FALSE, eval = TRUE, echo = FALSE}
library(knitr)
opts_chunk$set(
  fig.path = paste0("./figures/", purrr::pluck(params, "analysis_name", .default = FALSE)),
  fig.align = 'center',
  fig.show = 'asis',
  eval = TRUE,
  fig.width = 10,
  fig.height = 7,
  tidy = FALSE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  size = 'small',
  comment = "##",
  echo = TRUE,
  results = 'markup',
  dev = c("png", "pdf")
)
options(replace.assign = TRUE,
        width = 80)
```

```{r loadLibraries,results='hide'}
library("SummarizedExperiment")
library("ggplot2")
# library("limma")
library("pcaExplorer")
library("DESeq2")
library("GeneTonic")
library("DT")
library("org.Mm.eg.db") # or # library("org.Hs.eg.db")
library("biomaRt")
# library("clusterProfiler")
library("topGO")
library("pcaExplorer")
```

```{r load-functions}
# source("load_helper_functions.r")
devtools::load_all("/Users/edoardofilippi/Development/Projects/helper-functions-rbioinfo/")
# remotes::install_github("https://github.com/edo98811/helper-functions-rbioinfo.git@devel")
library("rbioinfoHelper")
```

# Importing files into R and preparing workspace

We start by importing all the necessary data.
These variables will be created in the workspace:

* anno_df
* anns
* gse object (if needed)
* experiment_metadata

If anno_df and anns still need to be created they will not be loaded
Note: gse must have at least an ensembl_gene_id column, the rownames must be either ensembl_gene_id or gene_ranges

```{r load-function}
prepare_workspace(params)
```

Add metadata to the gse object

```{r load-gse}
if (purrr::pluck(params, "load_gse", .default = FALSE)) colData(gse) <- S4Vectors::DataFrame(experiment_metadata)
```

If loading parameter set to true load the already analyzed data necessary objects

```{r load-data, eval = TRUE}
if (purrr::pluck(params, "load_dds_objects", .default = FALSE)) 
  load_analyses(params)
```

Subset using the funciton defined precently 

```{r subset-complete_dds, eval = TRUE}
if (
  purrr::pluck(params, "subset_gse", .default = FALSE) && 
  purrr::pluck(params, "load_gse", .default = TRUE)
  ) gse <- subset_gse(gse)
```
