---
title: "how_to"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how_to}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r packages}
library("macrophage")
library("DESeq2")
library("org.Hs.eg.db")
library("AnnotationDbi")
library("clusterProfiler")
library("limma")
library("edgeR")
library("DeeDeeExperiment")
library("GO.db")
library("topGO")
```

```{r setup}
library(rbioinfoHelper)
```

# Setting up parameters

Normally these parameters can be set in the yaml header of the Rmd file, making the analysis easy to reproduce.
```{r setup-par}
devtools::load_all()
params <- list(
  # data parameters, they can be also empty
  source_object = system.file("extdata", "gse.RDS", package = "rbioinfoHelper"),
  metadata_file = "", # we dont want to load metadata
  analysis_name = "example_analysis", # where the results will be saved, and from where they will be loaded

  # analysis parameters
  create_annotation_df = TRUE,
  run_computations = TRUE,
  save_results = TRUE,
  species = "Hs",

  # output parameters
  output_directory = "results/",
  output_name = "example_analysis"
)
stopifnot(check_params(params))
```


# Importing files into R and preparing workspace

We start by importing all the necessary data. We actually already ran the analysis adn the data is stored in ext/instdata. This means we can simply run the prepare workspace command. We also did not prepare the annotations.

```{r load-data}
prepare_workspace(params) # to set up the environment (load the starting object, metadata
gse <- source_object
load_results(params) # load results if we want to skip computations
```

The macrophage dataset requires BP ontology, we delete all the genes from the dataset that are not mapped in it.

```{r check-go}
rownames(gse) <- gsub("\\..*$", "", rownames(gse)) # To remove gene version

ontology_df <- AnnotationDbi::select(
  org.Hs.eg.db,
  keys = rownames(gse),
  columns = c("GO", "SYMBOL"),
  keytype = "ENSEMBL"
)

# remove rows with NA GO terms
ontology_df <- ontology_df[!is.na(ontology_df$GO), ]
ontology_df <- ontology_df[!is.na(ontology_df$SYMBOL), ]
ontology_df <- ontology_df[ontology_df$ONTOLOGY == "BP", ]

# remove duplicates for SYMBOL-GO pairs
ontology_df <- ontology_df[!duplicated(ontology_df[c("SYMBOL", "GO")]), ]

gse <- gse[rownames(gse) %in% ontology_df$ENSEMBL, ]
```
# Creating DeeDeeExperiment object

```{r create-dde}
# create DeeDeeExperiment object
dde <- DeeDeeExperiment(gse)
```

# DESeq2 analysis

```{r dds-creation}
# visualizedesign matrix
if (params$run_computations) {
  dds <- DESeqDataSet(gse, design = ~condition)

  # keep <- rowSums(counts(dds) >= 10) >= 6
  # dds <- dds[keep, ]

  dds <- DESeq(dds)
}
```

# Gene annotation

To create the gene annotation tables (if create_annotation_df = TRUE)

```{r gene-annotation, eval = TRUE}
create_annotations(params, rownames(dde))

# table(anns$SYMBOL[!is.na(anns$SYMBOL)] %in% keys(org.Hs.eg.db, keytype="SYMBOL"))
```

## Contrasts

```{r contrasts, eval = TRUE}
resultsNames(dds)
contrasts_list <- c(
  "condition_IFNg_vs_naive"
)
FDR <- 0.1
```

Here to plot the results, same thing if using another package (eg limma).

```{r de}
# Add all the
if (params$run_computations) {
  for (contrast_name in contrasts_list) {
    res <- results(dds, name = contrast_name, alpha = FDR)

    # Variance shrinkage
    res <- lfcShrink(dds, coef = contrast_name, res = res, type = "apeglm")

    dde <- addDEA(dde, res)
    dde <- renameDEA(dde, "res", contrast_name)
  }
}
```

```{r, echo = FALSE,include = FALSE, eval = T}
# You need this code to conduct the magic dependences attaching...
DT::datatable(matrix())
```

## Results
```{r de-tables, results = 'asis'}
for (contrast_name in contrasts_list) {
  table <- data.frame(DEA(dde, contrast_name, format = "original"))
  table <- dea_for_report(table, anns, interactive = TRUE, alpha = FDR)
  show_interactive_table(table, paste0("DE results results for ", contrast_name), knitting = TRUE)
  print(ideal::plot_ma(DEA(dde, contrast_name, format = "original"), ylim = c(-2, 2), title = contrast_name, FDR = FDR))
}
```


# Enrichment

## Analysis with topGO
```{r setup-enrichment}
expressed_ENSEMBL <- rownames(dds)[rowSums(assay(dds)) > 0]
expressed_SYMBOL <- anns$SYMBOL[match(expressed_ENSEMBL, anns$ENSEMBL)]
expressed_SYMBOL <- expressed_SYMBOL[!is.na(expressed_SYMBOL)]

# expressed_SYMBOL <- sample(keys(org.Hs.eg.db, keytype="SYMBOL"), 5000)
# de_genes <- sample(expressed_SYMBOL, 200)
```

```{r enrichment, eval = TRUE}
# enrichment with topGO
mapping <- "org.Hs.eg.db"

# enrichment with topGO
if (params$run_computations) {
  for (dea_name in names(getDEAList(dde))) {
    dea_table <- DEA(dde, dea_name, format = "original")
    dea_table <- dea_table[rowSums(is.na(dea_table)) == 0, ]
    if (!sum(dea_table$padj < FDR)) next # skip if no de
    dea_table <- dea_table[dea_table$padj < FDR, ]

    # Run topGO enrichment
    de_genes <- anns$SYMBOL[match(rownames(dea_table), anns$ENSEMBL)]
    topGO_results <- mosdef::run_topGO(
      de_genes = de_genes,
      bg_genes = expressed_SYMBOL,
      ontology = "BP",
      gene_id = "ENSEMBL",
      add_gene_to_terms = TRUE,
      mapping = mapping
    )

    # Add the fea to the
    dde <- addFEA(
      dde,
      topGO_results,
      de_name = dea_name
    )

    dde <- renameFEA(dde, "topGO_results", paste0(dea_name, "_fea"))
    dde <- linkDEAandFEA(dde, dea_name, paste0(dea_name, "_fea"), force = FALSE)
  }
}
```

```{r, echo = FALSE,include = FALSE}
# You need this code to conduct the magic dependences attaching...
DT::datatable(matrix())
```

## Results
```{r enrichment-tables, results = 'asis'}
for (contrast_fea in names(get_fea_list_report(dde))) {
  table <- FEA(dde, contrast_fea, format = "original")
  table <- fea_for_report(table, interactive = TRUE, alpha = FDR)
  show_interactive_table(table, paste0("Enrichment results for ", contrast_name), knitting = TRUE)
}
```


# Save 

```{r save, eval=TRUE}
save_results(
  params,
  list(dde = dde, anns = anns)
)
```


# Session Information {-}

This document has been generated with...

```{r}
sessionInfo()
```