---
title: >
  Exploratory Data Analysis 
author: >
  In vitro stimulation of Tubular Epithelial Cells Analysis </br>
  Najla Abassi and Federico Marini </br>
  University Medical Center - Mainz (Germany)
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
    code_folding: show
    number_sections: true
    theme: cosmo
date: "`r BiocStyle::doc_date()`" 
editor_options: 
  chunk_output_type: console
---

```{r echo = F}
options(error = function(){    # scream on error
  beepr::beep(9)
  Sys.sleep(1)
  }
)

.Last <- function() {          # Beep when knitting ends
  beepr::beep(8)
  Sys.sleep(7)
}

```

```{r setup, include=FALSE, cache=FALSE, eval = TRUE, echo = FALSE}
library(knitr)
opts_chunk$set(
  fig.path = paste0('./_figures_TEC_stim/'),
  fig.align = 'center',
  fig.show = 'asis',
  eval = TRUE,
  fig.width = 10,
  fig.height = 7,
  tidy = FALSE,
  message = FALSE,
  warning = FALSE,
  size = 'small',
  comment = "##",
  echo = TRUE,
  results = 'markup',
  dev = c("png", "pdf")
)
options(replace.assign = TRUE,
        width = 80)
```

```{r loadLibraries,results='hide'}
library("SummarizedExperiment")
library("ggplot2")
library("limma")
library("pcaExplorer")
library("DESeq2")
library("GeneTonic")
library("DT")
library("org.Mm.eg.db")
library("biomaRt")
library("clusterProfiler")
library("topGO")
library("plotly")
```

# Importing quantification files into R

We start by loading the `gse.RDS` object that we created containing the counts for all 32 samples.

```{r load-data}
gse <- readRDS("gse.RDS")

gse
```

```{r load-metadata}
# load metadata
coldata_all <- readxl::read_xlsx("sample_sheet_PTPRZ1_analysis.xlsx", col_names = T)

head(coldata_all)
```

# EDA - Exploratory Data Analysis : Role of PTPRZ1 in Acute Kidney Injury/ Chronic Kidney Disease

In vitro stimulation of cell culture (renal tubular epithelial cells TEC) was performed with LPS + IFNy.

We would like to know if PTPRZ1 promotes IC recruitment/activation and profibrotic signaling in murine TEC. For this we will subset only the PTPRZ1 Het and PTPRZ1 KO that were included in this experiment

```{r subset}
gse1 <- gse[, 13:32]

gse1

colnames(gse1)
```

Let's take a look at the experiment details:

```{r summary}
table(gse1$id) # how many samples
table(gse1$genotype) # how many sample per genotype
table(gse1$condition) # how many patients condition
table(gse1$group) # how many sample per group
table(gse1$sample_id) # how many sample
table(gse1$donor_animal) # how many sample per animal
```

We want to find the changes in gene expression that can be associated with knocking out PTPRZ1.

```{r set-factor}
colData(gse1)$id <- factor(colData(gse1)$id)
colData(gse1)$genotype <- factor(colData(gse1)$genotype)
colData(gse1)$condition <- factor(colData(gse1)$condition)
colData(gse1)$group <- factor(colData(gse1)$group)
colData(gse1)$sample_id <- factor(colData(gse1)$sample_id) 
colData(gse1)$donor_animal <- factor(colData(gse1)$donor_animal) 
```

Now let's create the `DESeqDataSet` object using the `RangedSummarizedExperiment` :

```{r dds}
dds <- DESeqDataSet(gse1, design = ~group)
dds
```

```{r set-levels}
lapply(colData(dds), levels)
```

```{r check-levels}
dds$id
dds$genotype
dds$condition
dds$group
dds$sample_id
dds$donor_animal
```

Let's how many genes have non-zero counts across all samples.

```{r unexp-genes}
nrow(dds) # total number of genes
table(rowSums(assay(dds, "counts")) == 0)
```

Let's check the number of assigned reads per sample as a proxy of the sequencing depth

```{r plot-seq-depth}
myd <- data.frame(
  counts = colSums(counts(dds)),
  sample = colnames(dds))
ggplot(myd, aes(x = sample,weight = counts, fill = sample)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme_bw() + coord_flip()
```

# PCA plots

```{r pca}
vst <- vst(dds)

pcaplot(vst, intgroup = "id", ellipse = FALSE, text_labels = FALSE, ntop = 5000)
pcaplot(vst, intgroup = "genotype", ellipse = FALSE, text_labels = FALSE, ntop = 5000)
pcaplot(vst, intgroup = "condition", ellipse = FALSE, text_labels = FALSE, ntop = 5000)
pcaplot(vst, intgroup = "group", ellipse = FALSE, text_labels = FALSE, ntop = 5000)
pcaplot(vst, intgroup = "sample_id", ellipse = FALSE, text_labels = FALSE, ntop = 5000)
pcaplot(vst, intgroup = "donor_animal", ellipse = FALSE, text_labels = FALSE, ntop = 5000)

```

The main driver of variability seems to be the `condition` , which means whether cells are stimulated or not

# Gene Annotation

```{r}
# creating an annotation df
# we need to perform some transformation to the ENSEMBL ids
head(rownames(rowData(dds)))
rowdata <- as.data.frame(rowData(dds))
row_names <- rownames(rowdata)
modified_row_names <- gsub("\\.[0-9]*$", "", row_names)
rownames(dds) <- modified_row_names
head(rownames(rowData(dds)))

anno_df <- pcaExplorer::get_annotation_orgdb(dds, "org.Mm.eg.db", "ENSEMBL")

```

```{r}
library("biomaRt")
mart <- useMart(biomart="ENSEMBL_MART_ENSEMBL", dataset="mmusculus_gene_ensembl")

anns <- getBM(attributes = c("ensembl_gene_id", "external_gene_name", "description"), 
              filters = "ensembl_gene_id",
              values = rownames(dds), 
              mart = mart)

anns <- anns[match(rownames(dds), anns$ensembl_gene_id), ]
# then this will be used to create the GeneTonicList later
```


# DE - Differential Expression, preliminary results

```{r}
# this is a collection of functions to make the results nicer
createLinkGO <- function(val) {
  sprintf('<a href="http://amigo.geneontology.org/amigo/term/%s" target="_blank" class="btn btn-primary">%s</a>',val,val)
}

createLinkENS  <- function(val, species="Mus_musculus") {
  paste0('<a href="http://www.ensembl.org/',species,'/Gene/Summary?g=',val,'" target="_blank" class="btn btn-primary">',val,'</a>')
}

createLinkGeneSymbol <- function(val) {
  # possibilities:
  # ncbi
  # genecards
  paste0('<a href="http://www.ncbi.nlm.nih.gov/gene/?term=',val,'[sym]" target="_blank" class="btn btn-primary">',val,'</a>')
}

alltheresults <- function(resuSet, dds_obj, contrast, FDR, anno_df, anns, species) {
  
  # id_contrast <- paste0(contrast[2],"_vs_",contrast[3])
  id_contrast <- contrast
  resuSet[[id_contrast]] <- list()
  
  # mycoef <- resultsNames(dds_obj)[2]
  mycoef <- contrast
  
  message("Extracting results...")
  resuSet[[id_contrast]][["res_DESeq"]] <- results(dds_obj,name = mycoef, alpha = FDR)
  # message("Performing LFC shrinkage...")
  resuSet[[id_contrast]][["res_DESeq"]] <- lfcShrink(dds_obj,coef = mycoef,res = resuSet[[id_contrast]][["res_DESeq"]], type = "apeglm") 
  resuSet[[id_contrast]][["res_DESeq"]]$SYMBOL <- anno_df$gene_name[match(rownames(resuSet[[id_contrast]][["res_DESeq"]]), anno_df$gene_id)]
  
  message("Summary MAplot...")
  summary(resuSet[[id_contrast]][["res_DESeq"]])
  resuSet[[id_contrast]][["maplot_res"]] <- 
    ideal::plot_ma(resuSet[[id_contrast]][["res_DESeq"]], ylim = c(-2,2), title = id_contrast, FDR = FDR)
  
  message("Extracting tables...")
  resuSet[[id_contrast]][["tbl_res_all"]] <- deseqresult2df(resuSet[[id_contrast]][["res_DESeq"]])
  resuSet[[id_contrast]][["tbl_res_all"]]$geneSymbol <- anno_df$gene_name[match(resuSet[[id_contrast]][["tbl_res_all"]]$id, anno_df$gene_id)]
  resuSet[[id_contrast]][["tbl_res_all"]]$description <- anns$description[match(resuSet[[id_contrast]][["tbl_res_all"]]$id, anns$ensembl_gene_id)]
  
  message("Extracting DEtables...")
  resuSet[[id_contrast]][["tbl_res_DE"]] <- deseqresult2df(resuSet[[id_contrast]][["res_DESeq"]],FDR = FDR)
  resuSet[[id_contrast]][["tbl_res_DE"]]$geneSymbol <- anno_df$gene_name[match(resuSet[[id_contrast]][["tbl_res_DE"]]$id, anno_df$gene_id)]
  resuSet[[id_contrast]][["tbl_res_DE"]]$description <- anns$description[match(resuSet[[id_contrast]][["tbl_res_DE"]]$id, anns$ensembl_gene_id)]
  resuSet[[id_contrast]][["tbl_res_DE"]]$chromosome_name <- anns$chromosome_name[match(resuSet[[id_contrast]][["tbl_res_DE"]]$id, anns$ensembl_gene_id)]
  
  if(nrow(resuSet[[id_contrast]][["tbl_res_DE"]]) > 0) {
    message("Generating interactive DEtable...")
    resuSet[[id_contrast]][["etbl_res_DE"]] <- resuSet[[id_contrast]][["tbl_res_DE"]]
    resuSet[[id_contrast]][["etbl_res_DE"]]$id <- createLinkENS(resuSet[[id_contrast]][["etbl_res_DE"]]$id, species = species)
    resuSet[[id_contrast]][["etbl_res_DE"]]$geneSymbol <- createLinkGeneSymbol(resuSet[[id_contrast]][["etbl_res_DE"]]$geneSymbol)
  }
  
  mybuttons <- c('copy', 'csv', 'excel', 'pdf', 'print')
  datatable(resuSet[[id_contrast]][["etbl_res_DE"]],caption = paste0(id_contrast,", DE genes"),escape=F, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = mybuttons))
  
  return(resuSet)
}
```

In this part we will compare the differences in gene expression within the `genotype`, while also accounting for the condition

```{r runde}
dds <- DESeq(dds)
```

These are our main contrasts

```{r contrast}
contrasts <- resultsNames(dds)
contrasts
```

```{r}
# set FDR 
FDR = 0.05
```

## DE - PTPRZ1 KO-ctr vs PTPRZ1 Het-ctrl

This would be PTPRZ1 KO-ctrl vs PTPRZ1 Het-ctrl (in other words here we are comparing KO vs WT when the cells are not stimulated)

```{r}
myresuSet <- list()
myresuSet <- alltheresults(
  myresuSet, dds,
  contrast = c("group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl"), FDR = FDR,
  anno_df = anno_df, anns = anns, species = "Mus_musculus")
myresuSet$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl$maplot_res
DT::datatable(myresuSet$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl$etbl_res_DE, escape = FALSE, rownames = FALSE)

```

```{r, echo=FALSE, eval=FALSE}
# DE - PTPRZ1 KO-stim vs PTPRZ1 Het-ctrl
# This would be PTPRZ1 KO-stim vs PTPRZ1 Het-ctrl
# myresuSet <- alltheresults(
#   myresuSet, dds,
#   contrast = c("group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_ctrl"), FDR = FDR,
#   anno_df = anno_df, anns = anns, species = "Mus_musculus")
# myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_ctrl$maplot_res
# DT::datatable(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_ctrl$etbl_res_DE, escape = FALSE, rownames = FALSE)
```

Now let's change the reference level

```{r relevel}
dds_releveled <- dds
dds_releveled$group <- relevel(dds_releveled$group, "PTPRZ_het_LPSIFNy")
dds_releveled <- nbinomWaldTest(dds_releveled)
resultsNames(dds_releveled)
```

```{r, echo=FALSE, eval=FALSE}
# DE - PTPRZ1 KO-ctrl vs PTPRZ 1 Het-stim
# This would be PTPRZ1 KO-ctrl vs PTPRZ 1 Het-stim
# myresuSet <- alltheresults(
#   myresuSet, dds_releveled,
#   contrast = c("group_PTPRZ_KO_ctrl_vs_PTPRZ_het_LPSIFNy"), FDR = FDR,
#   anno_df = anno_df, anns = anns, species = "Mus_musculus")
# myresuSet$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_LPSIFNy$maplot_res
# DT::datatable(myresuSet$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_LPSIFNy$etbl_res_DE, escape = FALSE, rownames = FALSE)
```


## DE - PTPRZ1 KO-stim vs PTPRZ 1 Het-stim

This would be PTPRZ1 KO-stim vs PTPRZ 1 Het-stim

```{r}
myresuSet <- alltheresults(
  myresuSet, dds_releveled,
  contrast = c("group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy"), FDR = FDR,
  anno_df = anno_df, anns = anns, species = "Mus_musculus")
myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy$maplot_res
DT::datatable(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy$etbl_res_DE, escape = FALSE, rownames = FALSE)
```

Now we would like to explore the differences due to the `condition`, while also accounting for the donor. The simplest approach is to subset by `genotype`, and then compare within each `genotype` the effect of the stimulation

```{r dds1}
# dds PTPRZ_het
dds_het <- dds[,dds$genotype == "PTPRZ_het"]
# drop the levels that are not in the PTPRZ_het group
dds_het$donor_animal <- droplevels(dds_het$donor_animal) 
dds_het$group <- droplevels(dds_het$group)

# dds PTPRZ_KO
dds_ko <- dds[,dds$genotype == "PTPRZ_KO"]
dds_ko$donor_animal <- droplevels(dds_ko$donor_animal)
dds_ko$group <- droplevels(dds_ko$group)
```

```{r recheck-levels}
dds_het$group
dds_het$donor_animal
dds_ko$group
dds_ko$donor_animal
```
Now we have to change the design to compare the effect of the stimulation. In other words, we will compare ctrl vs LPSIFNy:

```{r}
# design for WT group
design(dds_het) <- ~donor_animal + group
# design for KO group
design(dds_ko) <- ~donor_animal + group
```

```{r , echo=FALSE, eval=FALSE}
# dds_het_0 <- dds_het
# design(dds_het_0) <- ~group
# 
# dds_ko_0 <- dds_ko
# design(dds_ko_0) <- ~group
# 
# dds_het <- DESeq(dds_het)
# dds_het_0 <- DESeq(dds_het_0)
# 
# dds_ko <- DESeq(dds_ko)
# dds_ko_0 <- DESeq(dds_ko_0)
# 
# summary(results(dds_het), alpha = 0.05)
# summary(results(dds_het_0), alpha = 0.05)
# summary(results(dds_ko), alpha = 0.05)
# summary(results(dds_ko_0), alpha = 0.05)
```

## DE - PTPRZ1 Het-stim vs PTPRZ Het-ctrl

This would be PTPRZ1 Het-stim vs PTPRZ Het-ctrl

```{r runde1}
dds_het <- DESeq(dds_het)
```

These are our main contrasts

```{r}
contrasts <- resultsNames(dds_het)
contrasts
```

```{r}
myresuSet <- alltheresults(
  myresuSet, dds_het,
  contrast = c("group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl"), FDR = FDR,
  anno_df = anno_df, anns = anns, species = "Mus_musculus")
myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$maplot_res
DT::datatable(myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$etbl_res_DE, escape = FALSE, rownames = FALSE)
```

## DE - PTPRZ1 KO-stim vs PTPRZ KO-ctrl

This would be PTPRZ1 KO-stim vs PTPRZ KO-ctrl

```{r runde2}
dds_ko <- DESeq(dds_ko)
```

These are our main contrasts

```{r}
contrasts <- resultsNames(dds_ko)
contrasts
```

```{r}
myresuSet <- alltheresults(
  myresuSet, dds_ko,
  contrast = c("group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl"), FDR = FDR,
  anno_df = anno_df, anns = anns, species = "Mus_musculus")
myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$maplot_res
DT::datatable(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$etbl_res_DE, escape = FALSE, rownames = FALSE)
```

```{r}
de_het <- myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$tbl_res_DE
de_het$genotype <- "PTPRZ_het"
de_ko <- myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$tbl_res_DE
de_ko$genotype <- "PTPRZ_KO"

de <- rbind(de_het, de_ko)

# add a column of NAs
de$diffexpressed <- "NA"
# if log2Foldchange > 0 and pvalue < 0.05, set as "UP" 
de$diffexpressed[de$log2FoldChange > 0 & de$pvalue < 0.05] <- "UP"
# if log2Foldchange < 0 and pvalue < 0.05, set as "DOWN"
de$diffexpressed[de$log2FoldChange < 0 & de$pvalue < 0.05] <- "DOWN"

# add hover text
de$SYMBOL_genotype <- paste(de$SYMBOL,de$genotype,sep = "_")

# volcano plot of the stim vs unstim comparison
# when you hover the mouse you'll see 1 gene in the KO group and Het group
de %>%
  highlight_key(~id) %>%
  plotly::plot_ly(
    x = ~ log2FoldChange,
    y = ~ -log10(pvalue),
    type = "scatter",
    color = ~ diffexpressed,
    symbol = ~ genotype,
    symbols = c("circle", "square"),
    colors = c("#eb3642", "#3c5ade"),
    hoverinfo = "text",
    hovertext = ~ SYMBOL_genotype
  )  %>%
  plotly::highlight(on = "plotly_hover", off = "plotly_doubleclick") 

```


```{r}
res_het <- as.data.frame(myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$res_DESeq)
res_het$genotype <- "PTPRZ_het"
res_het$SYMBOL_genotype <- paste(res_het$SYMBOL,res_het$genotype,sep = "_")
res_het$id <- rownames(res_het)

res_ko <- as.data.frame(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$res_DESeq)
res_ko$genotype <- "PTPRZ_KO"
res_ko$SYMBOL_genotype <- paste(res_ko$SYMBOL,res_ko$genotype,sep = "_")
res_ko$id <- rownames(res_ko)

res_total <- merge(
  res_het[,c("log2FoldChange","pvalue","padj","SYMBOL","genotype","SYMBOL_genotype","id")],
  res_ko[,c("log2FoldChange","pvalue","padj","genotype","SYMBOL_genotype","id")],
  by= "id",
  all=TRUE,
  suffixes = c("_het", "_ko")
)

# # add a column of NAs
# res_total$diffexpressed <- "NA"
# # if both log2Foldchange > 0 and pvalue < 0.05, set as "HET_UP-KO_UP" 
# res_total$diffexpressed[res_total$log2FoldChange_het > 0 & res_total$log2FoldChange_ko > 0 & res_total$pvalue_het < 0.05 & res_total$pvalue_ko < 0.05] <- "HET_UP-KO_UP"
# # if both log2Foldchange < 0 and pvalue < 0.05, set as "HET_DOWN-KO_DOWN"
# res_total$diffexpressed[res_total$log2FoldChange_het > 0 & res_total$log2FoldChange_ko > 0 & res_total$pvalue_het < 0.05 & res_total$pvalue_ko < 0.05] <- "HET_DOWN-KO_DOWN"
# # if log2FoldChange_het > 0 and log2FoldChange_ko < 0 and pvalue < 0.05, set as "HET_UP-KO_DOWN"
# res_total$diffexpressed[res_total$log2FoldChange_het > 0 & res_total$log2FoldChange_ko < 0 & res_total$pvalue_het < 0.05 & res_total$pvalue_ko < 0.05] <- "HET_UP-KO_DOWN"
# # if log2FoldChange_het < 0 and log2FoldChange_ko > 0 and pvalue < 0.05, set as "KO_UP-HET_DOWN"
# res_total$diffexpressed[res_total$log2FoldChange_het < 0 & res_total$log2FoldChange_ko > 0 & res_total$pvalue_het < 0.05 & res_total$pvalue_ko < 0.05] <- "KO_UP-HET_DOWN"
# 
# ## should we drop NA?
# # res_total <- tidyr::drop_na(res_total)
# # NA are plotted as white?
# 
# 
# plotly::plot_ly(
#   res_total,
#     x = ~ log2FoldChange_het,
#     y = ~ log2FoldChange_ko,
#     type = "scatter",
#     color = ~ diffexpressed,
#     colors = c("#eb3642", "#3c5ade" , "#4dcdd6", "#3be395"),
#     hoverinfo = "text",
#   hovertext = ~ SYMBOL
#   )


```

```{r}
# add a column of NAs
res_total$diffexpressed <- "NA"
# if both log2Foldchange > 0, set as "HET_UP-KO_UP" 
res_total$diffexpressed[res_total$log2FoldChange_het > 0 & res_total$log2FoldChange_ko > 0] <- "HET_UP-KO_UP"
# if both log2Foldchange < 0, set as "HET_DOWN-KO_DOWN"
res_total$diffexpressed[res_total$log2FoldChange_het < 0 & res_total$log2FoldChange_ko < 0] <- "HET_DOWN-KO_DOWN"
# if log2FoldChange_het > 0 and log2FoldChange_ko < 0, set as "HET_UP-KO_DOWN"
res_total$diffexpressed[res_total$log2FoldChange_het > 0 & res_total$log2FoldChange_ko < 0] <- "HET_UP-KO_DOWN"
# if log2FoldChange_het < 0 and log2FoldChange_ko > 0, set as "KO_UP-HET_DOWN"
res_total$diffexpressed[res_total$log2FoldChange_het < 0 & res_total$log2FoldChange_ko > 0] <- "KO_UP-HET_DOWN"

plotly::plot_ly(
  res_total,
    x = ~ log2FoldChange_het,
    y = ~ log2FoldChange_ko,
    type = "scatter",
    color = ~ diffexpressed,
    colors = c("#eb3642", "#3c5ade" , "#4dcdd6", "#3be395"),
    hoverinfo = "text",
  hovertext = ~ SYMBOL
  )
```

```{r, eval=FALSE}
# save "DE" tables in an excel sheet
library("openxlsx")
library("stringr")

# Create a new workbook
wb <- createWorkbook()

# Iterate over each element in the `myresuSet` and add a new sheet
for (contrast in names(myresuSet)) {
  truncked_name <- str_remove_all(contrast,"group_|PTPRZ_")
  addWorksheet(wb, sheetName = truncked_name)
  writeData(wb, sheet = truncked_name, x = myresuSet[[contrast]]$tbl_res_DE)
}

saveWorkbook(wb, "DE_gene_list.xlsx", overwrite = TRUE)
```

# Functional Enrichment Analysis

Now we would like to determine which biological pathways are associated with these DEGs. This will help us understand the processes and functions these genes are involved in.

```{r}
expressedInAssay <- (rowSums(assay(dds))>0)
geneUniverseExprENS <- rownames(dds)[expressedInAssay]
geneUniverseExpr <- anno_df$gene_name[match(geneUniverseExprENS, anno_df$gene_id)]
```

## Analysis with clusterProfiler

```{r}
# enrichment with clusterProfiler

for(i in names(myresuSet)) {
  message(i)
  if(nrow(myresuSet[[i]][["tbl_res_DE"]]) > 0) {
    myresuSet[[i]][["clupro_tbl"]] <- 
      enrichGO(
        gene = myresuSet[[i]][["tbl_res_DE"]]$geneSymbol,
        universe      = geneUniverseExpr,
        keyType       = "SYMBOL",
        OrgDb         = org.Mm.eg.db,
        ont           = "BP",
        pAdjustMethod = "BH",
        pvalueCutoff  = 0.01,
        qvalueCutoff  = 0.02,
        readable      = FALSE)
  }
}
```


### Enrichment results for PTPRZ KO-ctrl vs PTPRZ het-ctrl

```{r}
DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl$clupro_tbl), rownames = FALSE)
```

```{r, echo=FALSE,eval=FALSE}
# display enrich results of group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_ctrl
# DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_ctrl$clupro_tbl), rownames = FALSE)
```

```{r, echo=FALSE,eval=FALSE}
# display enrich results of group_PTPRZ_KO_ctrl_vs_PTPRZ_het_LPSIFNy
# DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_LPSIFNy$clupro_tbl), rownames = FALSE)
```

### Enrichment results for PTPRZ KO-LPSIFNy vs PTPRZ het-LPSIFNy

```{r}
DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy$clupro_tbl), rownames = FALSE)
```

### Enrichment results for PTPRZ het-LPSIFNy vs PTPRZ het-ctrl

```{r}
DT::datatable(as.data.frame(myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$clupro_tbl), rownames = FALSE)
```

### Enrichment results for PTPRZ-KO LPSIFNy vs PTPRZ KO-ctrl

```{r}
DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$clupro_tbl), rownames = FALSE)
```
```{r, eval=FALSE}
# save "Enrich" tables in an excel sheet
# Create a new workbook
wb <- createWorkbook()

# Iterate over each element in the `myresuSet` and add a new sheet
for (contrast in names(myresuSet)) {
  truncked_name <- str_remove_all(contrast,"group_|PTPRZ_")
  addWorksheet(wb, sheetName = truncked_name)
  writeData(wb, sheet = truncked_name, x = myresuSet[[contrast]]$clupro_tbl)
}

saveWorkbook(wb, "Enrich_results_ClusterPro.xlsx", overwrite = TRUE)
```

## Analysis with topGO

```{r}
# enrichment with topGO
for (i in names(myresuSet)) {
  message(i)
  if (nrow(myresuSet[[i]][["tbl_res_DE"]]) > 0) {
  myresuSet[[i]][["topGO_tbl"]] <- 
    topGOtable(DEgenes = myresuSet[[i]][["tbl_res_DE"]]$geneSymbol, 
               BGgenes = geneUniverseExpr, 
               ontology = "BP",
               geneID = "symbol",
               addGeneToTerms = TRUE, 
               topTablerows = 5000,
               mapping = "org.Mm.eg.db")
  }
}
```


### Enrichment results for PTPRZ KO-ctrl vs PTPRZ het-ctrl

```{r}
DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl$topGO_tbl), rownames = FALSE)
```

```{r, echo=FALSE,eval=FALSE}
# enrich results for group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_ctrl
# DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_ctrl$topGO_tbl), rownames = FALSE)
```

```{r}
# enrich results for group_PTPRZ_KO_ctrl_vs_PTPRZ_het_LPSIFNy
DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_LPSIFNy$topGO_tbl), rownames = FALSE)
```

### Enrichment results for PTPRZ KO-LPSIFNy vs PTPRZ het-LPSIFNy

```{r}
DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy$topGO_tbl), rownames = FALSE)
```

### Enrichment results for PTPRZ het-LPSIFNy vs PTPRZhet-ctrl

```{r}
DT::datatable(as.data.frame(myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$topGO_tbl), rownames = FALSE)
```

### Enrichment results for PTPRZ-KO LPSIFNy vs PTPRZ KO-ctrl

```{r}
DT::datatable(as.data.frame(myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$topGO_tbl), rownames = FALSE)
```

```{r, eval=FALSE}
# save "Enrich" tables in an excel sheet
# Create a new workbook
wb <- createWorkbook()

# Iterate over each element in the `myresuSet` and add a new sheet
for (contrast in names(myresuSet)) {
  truncked_name <- str_remove_all(contrast,"group_|PTPRZ_")
  addWorksheet(wb, sheetName = truncked_name)
  writeData(wb, sheet = truncked_name, x = myresuSet[[contrast]]$topGO_tbl)
}

saveWorkbook(wb, "Enrich_results_TopGO.xlsx", overwrite = TRUE)
```

# Interactive Exploration with `GeneTonic`

This data can be further explored interactively with `r BiocStyle::Biocpkg("GeneTonic")`. For this, we need to create a `GeneTonicList`

```{r}
# function to easily generate GeneTonicLists for different contrasts
resuset_to_gtl <- function(myresuSet, result_name, dds, anno_df, enrich_algo) {
  dds_gtl <- dds
  anno_df_gtl <- anno_df
  res_de_gtl <- myresuSet[[result_name]][["res_DESeq"]]
  
  if (enrich_algo == "topGO_tbl") {
    res_enrich_gtl <- GeneTonic::shake_topGOtableResult(myresuSet[[result_name]][["topGO_tbl"]])
  } else if (enrich_algo == "clupro_tbl") {
    res_enrich_gtl <- GeneTonic::shake_enrichResult(myresuSet[[result_name]][["clupro_tbl"]])
  }
  
  
  gtl_assembled <- GeneTonicList(
    dds = dds_gtl,
    res_de = res_de_gtl,
    res_enrich = res_enrich_gtl,
    annotation_obj = anno_df_gtl
  )
  
  return(gtl_assembled)
}
```

```{r}
gtl_KO_unstim_vs_HET_unstim <- resuset_to_gtl(myresuSet, "group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl", dds, anno_df, enrich_algo = "clupro_tbl")
gtl_KO_stim_vs_HET_stim <- resuset_to_gtl(myresuSet, "group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy", dds_releveled, anno_df, enrich_algo = "clupro_tbl")
gtl_HET_stim_vs_HET_unstim <- resuset_to_gtl(myresuSet, "group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl", dds, anno_df, enrich_algo = "clupro_tbl")
gtl_KO_stim_vs_KO_unstim <- resuset_to_gtl(myresuSet, "group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl", dds_releveled, anno_df, enrich_algo = "clupro_tbl")

```

```{r save1, eval=FALSE}
saveRDS(gtl_KO_unstim_vs_HET_unstim, "gtl_KO_unstim_vs_HET_unstim.RDS")
saveRDS(gtl_KO_stim_vs_HET_stim, "gtl_KO_stim_vs_HET_stim.RDS")
saveRDS(gtl_HET_stim_vs_HET_unstim, "gtl_HET_stim_vs_HET_unstim.RDS")
saveRDS(gtl_KO_stim_vs_KO_unstim,"gtl_KO_stim_vs_KO_unstim.RDS")
```

```{r save, eval=FALSE}
saveRDS(dds, "dds_tecs_ptprz_ko_vs_het.RDS")
saveRDS(dds_het,"dds_het_stim_vs_unstim.RDS")
saveRDS(dds_ko,"dds_ko_stim_vs_unstim.RDS")
```

Now  we can use `GeneTonic` to explore the data by simply typing …

```{r eval=FALSE}
gtl <- readRDS("_gtl/gtl_HET_stim_vs_HET_unstim.RDS") # just import any GeneTonicList
GeneTonic(gtl = gtl)
```

# Further Analysis ...

## DE Analysis without LFC shrinkage 

```{r, echo=FALSE}
alltheresults <- function(resuSet, dds_obj, contrast, FDR, anno_df, anns, species) {
  
  # id_contrast <- paste0(contrast[2],"_vs_",contrast[3])
  id_contrast <- contrast
  resuSet[[id_contrast]] <- list()
  
  # mycoef <- resultsNames(dds_obj)[2]
  mycoef <- contrast
  
  message("Extracting results...")
  resuSet[[id_contrast]][["res_DESeq"]] <- results(dds_obj,name = mycoef, alpha = FDR)
  # message("Performing LFC shrinkage...")
  #resuSet[[id_contrast]][["res_DESeq"]] <- lfcShrink(dds_obj,coef = mycoef,res = resuSet[[id_contrast]][["res_DESeq"]], type = "apeglm") 
  resuSet[[id_contrast]][["res_DESeq"]]$SYMBOL <- anno_df$gene_name[match(rownames(resuSet[[id_contrast]][["res_DESeq"]]), anno_df$gene_id)]
  
  message("Summary MAplot...")
  summary(resuSet[[id_contrast]][["res_DESeq"]])
  resuSet[[id_contrast]][["maplot_res"]] <- 
    ideal::plot_ma(resuSet[[id_contrast]][["res_DESeq"]], ylim = c(-2,2), title = id_contrast, FDR = FDR)
  
  message("Extracting tables...")
  resuSet[[id_contrast]][["tbl_res_all"]] <- deseqresult2df(resuSet[[id_contrast]][["res_DESeq"]])
  resuSet[[id_contrast]][["tbl_res_all"]]$geneSymbol <- anno_df$gene_name[match(resuSet[[id_contrast]][["tbl_res_all"]]$id, anno_df$gene_id)]
  resuSet[[id_contrast]][["tbl_res_all"]]$description <- anns$description[match(resuSet[[id_contrast]][["tbl_res_all"]]$id, anns$ensembl_gene_id)]
  
  message("Extracting DEtables...")
  resuSet[[id_contrast]][["tbl_res_DE"]] <- deseqresult2df(resuSet[[id_contrast]][["res_DESeq"]],FDR = FDR)
  resuSet[[id_contrast]][["tbl_res_DE"]]$geneSymbol <- anno_df$gene_name[match(resuSet[[id_contrast]][["tbl_res_DE"]]$id, anno_df$gene_id)]
  resuSet[[id_contrast]][["tbl_res_DE"]]$description <- anns$description[match(resuSet[[id_contrast]][["tbl_res_DE"]]$id, anns$ensembl_gene_id)]
  resuSet[[id_contrast]][["tbl_res_DE"]]$chromosome_name <- anns$chromosome_name[match(resuSet[[id_contrast]][["tbl_res_DE"]]$id, anns$ensembl_gene_id)]
  
  if(nrow(resuSet[[id_contrast]][["tbl_res_DE"]]) > 0) {
    message("Generating interactive DEtable...")
    resuSet[[id_contrast]][["etbl_res_DE"]] <- resuSet[[id_contrast]][["tbl_res_DE"]]
    resuSet[[id_contrast]][["etbl_res_DE"]]$id <- createLinkENS(resuSet[[id_contrast]][["etbl_res_DE"]]$id, species = species)
    resuSet[[id_contrast]][["etbl_res_DE"]]$geneSymbol <- createLinkGeneSymbol(resuSet[[id_contrast]][["etbl_res_DE"]]$geneSymbol)
  }
  
  mybuttons <- c('copy', 'csv', 'excel', 'pdf', 'print')
  datatable(resuSet[[id_contrast]][["etbl_res_DE"]],caption = paste0(id_contrast,", DE genes"),escape=F, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = mybuttons))
  
  return(resuSet)
}
```


These are our main contrasts

```{r contrast_unshrunk}
dds_unshrunk <- dds
contrasts <- resultsNames(dds_unshrunk)
contrasts
```

### DE - PTPRZ1 KO-ctr vs PTPRZ1 Het-ctrl

This would be PTPRZ1 KO-ctrl vs PTPRZ1 Het-ctrl (in other words here we are comparing KO vs WT when the cells are not stimulated)

```{r}
myresuSet_unshrunk <- list()
myresuSet_unshrunk <- alltheresults(
  myresuSet_unshrunk, dds_unshrunk,
  contrast = c("group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl"), FDR = FDR,
  anno_df = anno_df, anns = anns, species = "Mus_musculus")
myresuSet_unshrunk$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl$maplot_res
DT::datatable(myresuSet_unshrunk$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl$etbl_res_DE, escape = FALSE, rownames = FALSE)

```

Now let's change the reference level

```{r}
dds_releveled_unshrunk <- dds_unshrunk
dds_releveled_unshrunk$group <- relevel(dds_releveled_unshrunk$group, "PTPRZ_het_LPSIFNy")
dds_releveled_unshrunk <- nbinomWaldTest(dds_releveled_unshrunk)
resultsNames(dds_releveled_unshrunk)
```

### DE - PTPRZ1 KO-stim vs PTPRZ 1 Het-stim

This would be PTPRZ1 KO-stim vs PTPRZ 1 Het-stim

```{r}
myresuSet_unshrunk <- alltheresults(
  myresuSet_unshrunk, dds_releveled_unshrunk,
  contrast = c("group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy"), FDR = FDR,
  anno_df = anno_df, anns = anns, species = "Mus_musculus")
myresuSet_unshrunk$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy$maplot_res
DT::datatable(myresuSet_unshrunk$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy$etbl_res_DE, escape = FALSE, rownames = FALSE)
```

### DE - PTPRZ1 Het-stim vs PTPRZ Het-ctrl

This would be PTPRZ1 Het-stim vs PTPRZ Het-ctrl

These are our main contrasts

```{r}
dds_het_unshrunk <- dds_het
contrasts <- resultsNames(dds_het_unshrunk)
contrasts
```

```{r}
myresuSet_unshrunk <- alltheresults(
  myresuSet_unshrunk, dds_het_unshrunk,
  contrast = c("group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl"), FDR = FDR,
  anno_df = anno_df, anns = anns, species = "Mus_musculus")
myresuSet_unshrunk$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$maplot_res
DT::datatable(myresuSet_unshrunk$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$etbl_res_DE, escape = FALSE, rownames = FALSE)
```

### DE - PTPRZ1 KO-stim vs PTPRZ KO-ctrl

This would be PTPRZ1 KO-stim vs PTPRZ KO-ctrl

These are our main contrasts

```{r}
dds_ko_unshrunk <- dds_ko
contrasts <- resultsNames(dds_ko_unshrunk)
contrasts
```

```{r}
myresuSet_unshrunk <- alltheresults(
  myresuSet_unshrunk, dds_ko_unshrunk,
  contrast = c("group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl"), FDR = FDR,
  anno_df = anno_df, anns = anns, species = "Mus_musculus")
myresuSet_unshrunk$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$maplot_res
DT::datatable(myresuSet_unshrunk$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$etbl_res_DE, escape = FALSE, rownames = FALSE)
```

```{r, eval=FALSE}
# save "DE" tables in an excel sheet
# Create a new workbook
wb <- createWorkbook()

# Iterate over each element in the `myresuSet` and add a new sheet
for (contrast in names(myresuSet_unshrunk)) {
  truncked_name <- str_remove_all(contrast,"group_|PTPRZ_")
  addWorksheet(wb, sheetName = truncked_name)
  writeData(wb, sheet = truncked_name, x = myresuSet_unshrunk[[contrast]]$tbl_res_DE)
}

saveWorkbook(wb, "DE_unshrunk_gene_list.xlsx", overwrite = TRUE)
```


## Visualizations ...

### Heatmap of the common DE genes in the comparison KO vs HET

```{r}
# heatmap
# start by creating gene signatures, which is the union of DE genes in ctrl and stim condition
de_ctr <- myresuSet$group_PTPRZ_KO_ctrl_vs_PTPRZ_het_ctrl$tbl_res_DE
de_stim <- myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_het_LPSIFNy$tbl_res_DE
# gene signature is the union
gene_signature <- rbind(de_ctr,de_stim)
# use the normalized counts
vst <- vst(dds)

vst_corrected <- vst
# vst_corrected <- vst[, vst_corrected$genotype == "PTPRZ_KO"]
# vst_corrected$donor_animal <- droplevels(vst_corrected$donor_animal)

# correct for batch effect 
assay(vst_corrected) <- limma::removeBatchEffect(assay(vst_corrected), vst_corrected$donor_animal)
#
vst <- vst_corrected

# create matrix for heatmap
de_union <- assay(vst)[unique(gene_signature$id),]
#rownames(de_union) <- gene_signature$SYMBOL
rownames(de_union) <- anno_df[rownames(de_union),]$gene_name
# create annot_col
df <- data.frame("condition" = vst$condition,
                 "genotype" = vst$genotype,
                 "donor" = vst$donor_animal)

rownames(df) <- vst$id

# reorder the columns for better visualization
new_order <- c(
  "sample_109", 
  "sample_183", 
  "sample_216", 
  "sample_249", 
  "sample_305",
  
  "sample_110", 
  "sample_184", 
  "sample_215",
  "sample_248", 
  "sample_304", 
  
  "sample_202", 
  "sample_353", 
  "sample_213", 
  "sample_241",
  "sample_363",

  "sample_201", 
  "sample_354", 
  "sample_222", 
  "sample_242", 
  "sample_362"
  
)
de_union <- de_union[, new_order]
# df <- df[order(df$genotype,df$condition), ]
df <- df[order(df$genotype,df$condition), ]
rownames(df) <- new_order


ann_colors = list(
    condition = c("ctrl" = "#8ee6ba", "LPSIFNy" = "#fb9a99"),
    genotype = c("PTPRZ_het" = "#478ffc", "PTPRZ_KO" = "#95d9f0"),
    donor = c("mouse_1" =  "#9e0142",
              "mouse_2" =  "#ff99c4",
              "mouse_3" =  "#f46d43",
              "mouse_4" =  "#fdae61",
              "mouse_5" =  "#fee08b",
              "mouse_6" =  "#e8e866",
              "mouse_7" =  "#bccf5b",
              "mouse_8" =  "#99ffe6",
              "mouse_9" =  "#66c2a5",
              "mouse_10" = "#cab2d6")
    
)


pheatmap::pheatmap((de_union - rowMeans(de_union))/rowSds(de_union),
                   # scale = "row",
                   border_color = "white",
                   annotation = df,
                   cluster_cols = F,
                   annotation_colors = ann_colors,
                   show_colnames = F)
```

### Figure 7E

```{r}
######################
# Figure 7E manuscript
######################
gene_signature_TEC <- c(
  "Clec7a", 
  "Clec4e", "Clec4n", "Il1a", "Il1b",
  "Il6", "Il12a", "Il12b", "Il34", "Il36b",
  "Tnf", "Ccl2", "Ccl3", "Ccl5", "Ccl7",
  "Ccl8", "Cxcl3", "Cxcl5","Cxcl9", "Ccl20",
  "Cxcl10", "Cxcl16", "Cx3cl1", "Cx3cr1", "Csf1",
  "Nos1", "Nos2", "Nos3", "Nox1", "Casp1",
  "Icam1", "Vcam1"
)

# Clec7e doesn't exist, instead we put Clec7a

# create matrix for heatmap
matched_id <- rowData(vst)$SYMBOL %in% gene_signature_TEC 
heatmap_data <- assay(vst)[matched_id, ]
rownames(heatmap_data) <- rowData(vst)$SYMBOL[matched_id]

# create annot_col
df <- data.frame("group" = vst$group)
rownames(df) <- vst$id

# reorder the columns for better visualization
# new_order <- c(
#   "sample_109", "sample_183", "sample_216", "sample_249", "sample_305",
#   "sample_110", "sample_184", "sample_215", "sample_248", "sample_304",
#   "sample_202", "sample_353", "sample_213", "sample_241", "sample_363",
#   "sample_201", "sample_354", "sample_222", "sample_242", "sample_362"
# )

heatmap_data <- heatmap_data[, new_order]
df <- as.data.frame(df[order(df$group),])
#rename col
names(df)[names(df) == "df[order(df$group), ]"] <- "group"
rownames(df) <- new_order

pheatmap::pheatmap(heatmap_data - rowMeans(heatmap_data),
                   annotation = df,
                    cluster_cols = F,
                   scale = "row",
                   border_color = "white",
                   show_colnames = F)

#  another alternative for the same plot
gs_heatmap(se=vst[,new_order],
           genelist = anno_df$gene_id[match(gene_signature_TEC, anno_df$gene_name)],
           #geneset_id = "GO:0002833",
           annotation_obj = anno_df,
           gtl = gtl_KO_stim_vs_KO_unstim,
           scale_row = T,
           show_row_dend = F,
           show_column_names = F,
           #row_km = 2,
           anno_col_info = "group")



```

you can choose any of the above plots :)

### Figure 11A - supplement

```{r}
#########################
# Figure 11A - supplement
#########################
total_results <- res_total

highlight_genes <- c("Ifi209", "Csf2", "Ccl20", "Ccl3", "Il1rn", "Tnfrsf1b")
# gene_colors <- c("Ifi209" = "#198038",
#                  "Csf2" = "#b28600",
#                  "Ccl20" = "#ee538b",
#                  "Ccl3" = "#ff0008",
#                  "Il1rn" = "#002d9c",
#                  "Tnfrsf1b" = "#1192e8")

total_results$highlight_color <- ifelse(total_results$SYMBOL %in% highlight_genes, 
                                    "red", 
                                    "gray")

# hide all the labels
total_results$SYMBOL <- ""
ix_label <- which(res_total$SYMBOL %in% highlight_genes)
# label only heilight_genes
total_results$SYMBOL[ix_label] <- res_total$SYMBOL[ix_label]


ggplot(total_results,
       aes(x = log2FoldChange_het, y = log2FoldChange_ko, label = SYMBOL)) +
  geom_point(
    aes(color = highlight_color),
    size = 1.3,
    alpha = ifelse(total_results$SYMBOL %in% highlight_genes, 1, 0.2)
  ) +
  scale_color_identity() +
  theme_linedraw() +
  ggrepel::geom_text_repel(box.padding = 0.5, max.overlaps = Inf) +
  geom_hline(yintercept = 0,
             color = "grey",
             linetype = "dashed") +
  geom_vline(xintercept = 0,
             color = "grey",
             linetype = "dashed")

```

### Figure 11B - supplement

```{r}
#########################
# Figure 11B - supplement
#########################

# add enrich results
enrich_res_ko <- myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$topGO_tbl
enrich_res_het <- myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$topGO_tbl

# calculate z score
enrich_res_ko <- get_aggrscores(
  shake_topGOtableResult(enrich_res_ko),
  myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$res_DESeq,
  annotation_obj = anno_df
)

enrich_res_het <- get_aggrscores(
  shake_topGOtableResult(enrich_res_het),
  myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$res_DESeq,
  annotation_obj = anno_df
)


# add enrich results
# enrich_res_ko <- myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$clupro_tbl
# enrich_res_het <- myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$clupro_tbl
# 
# # extract the results (to obtain the z score)
# enrich_res_ko_df <- myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$clupro_tbl@result
# enrich_res_het_df <- myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$clupro_tbl@result

# shake results to be able to directly use it in GeneTonic
# enrich_res_ko_shaked <- shake_enrichResult(enrich_res_ko)
# enrich_res_het_shaked <- shake_enrichResult(enrich_res_het)
# 
# # add the zscores
# enrich_res_het_shaked$z_score <- enrich_res_het_df$zScore
# enrich_res_ko_shaked$z_score <- enrich_res_ko_df$zScore

# subset for the terms of interest 
go_terms <- c(
  "GO:0001819",
  "GO:0031349",
  "GO:0045088",
  "GO:0048525",
  "GO:0002250",
  "GO:0002697",
  "GO:0030198"
)

res_enrich1 <- enrich_res_ko[enrich_res_ko$gs_id %in% go_terms,]
res_enrich2 <- enrich_res_het[enrich_res_het$gs_id %in% go_terms,]

GeneTonic::gs_summary_overview_pair(res_enrich = res_enrich1,
                                    res_enrich2 = res_enrich2)


# gs_heatmap(se = vst[,new_order],
#            res_enrich = res_enrich1,
#            annotation_obj = anno_df,
#            geneset_id = go_terms[1],
#            anno_col_info = c("donor_animal","genotype","group"),
#            scale_row = T)

# heatmap for each pathway
for (i in go_terms) {
  p <- gs_heatmap(se = vst[,new_order],
           res_enrich = res_enrich1,
           annotation_obj = anno_df,
           geneset_id = i,
           anno_col_info = c("donor_animal","genotype","group"),
           scale_row = T,
           show_column_names = F
           )
  p
}

```

Or even better a heatmap with all the pathway together...

#### Heatmap pathway for the HET genotype

```{r}
# a heatmap with all the pathway together
scores_mat_het <- gs_scores(
  se = vst,
  res_de = myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$res_DESeq,
  res_enrich = enrich_res_het,
  annotation_obj = anno_df
)
#subset for terms of interest
go_terms <- c(
"positive regulation of cytokine production|GO:0001819",
"positive regulation of defense response|GO:0031349",
"regulation of innate immune response|GO:0045088",
"negative regulation of viral process|GO:0048525",
"adaptive immune response|GO:0002250",
"regulation of immune effector process|GO:0002697",
"extracellular matrix organization|GO:0030198"
)
####
scores_mat_subset_het <- scores_mat_het[rownames(scores_mat_het) %in% go_terms, ]

gs_scoresheat(scores_mat_subset_het,
              gs_ids = go_terms,
              n_gs = 10)

# topGO
# res_enrich_ko <- myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$topGO_tbl
# res_enrich_het <- myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$topGO_tbl
# enrich_res_ko_shaked <- shake_topGOtableResult(enrich_res_ko) # not calculated with pcaExplorer::topGOtable
# res_de_ko <- myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$res_DESeq
# res_de_het <- myresuSet$group_PTPRZ_het_LPSIFNy_vs_PTPRZ_het_ctrl$res_DESeq
# #calculate the z score
# z_scores <- get_aggrscores(res_enrich_ko, res_de_ko, anno_df)

```

#### Heatmap pathway for the KO genotype

```{r}
scores_mat_ko <- gs_scores(
  se = vst,
  res_de = myresuSet$group_PTPRZ_KO_LPSIFNy_vs_PTPRZ_KO_ctrl$res_DESeq,
  res_enrich = enrich_res_ko,
  annotation_obj = anno_df
)
#subset for terms of interest
go_terms <- c(
"positive regulation of cytokine production|GO:0001819",
"positive regulation of defense response|GO:0031349",
"regulation of innate immune response|GO:0045088",
"negative regulation of viral process|GO:0048525",
"adaptive immune response|GO:0002250",
"regulation of immune effector process|GO:0002697",
"extracellular matrix organization|GO:0030198"
)
####
scores_mat_subset_ko <- scores_mat_ko[rownames(scores_mat_ko) %in% go_terms, ]

gs_scoresheat(scores_mat_subset_ko,
              gs_ids = go_terms,
              n_gs = 10)
```



# Session information {-}

```{r sessioninfo}
BIUMmisc::sessioninfo_extra()
```
